Computer Organization Exam Review

### 2016-01-02 computer_organization/review.tex
update several

\begin{document}
\begin{enumerate}
\item $CPU\ Time = \frac{Clock\ Cycles}{Clock\ Rate}$
\item $CPU\ Time = Instructions \times \frac{Cycles}{Instruction} \times \frac{Seconds}{Cycle}$
\item $CPU\ Time = Clock\ Cycles \times Cycle\ Time$
\item $CPI = \frac{Clock\ Cycles}{Instructions}$
\item $CPU\ Time = Instruction\ Count \times CPI \times Cycle\ Time$
\item $CPU\ Time = Instruction\ Count \times \frac{CPI}{Clock\ Rate}$
\item $(Clock\ Cycles)_{Total} = \sum_{i = 1}^{n}\left(CPI_{i} \times Count_{i}\right)$
\item $MIPS = \frac{Instruction\ Count}{CPU\ Time \times 10^{6}}$
\item $Power = Capacitive\ Load \times (Voltage)^{2} \times Frequency$
\item $AMAT = Hit\ Time + Miss\ Rate \times Miss\ Penalty$
\end{enumerate}
\hrule
\begin{enumerate}
\item Identify the data hazards in the following code, and fix them using $nop$ (software).
\begin{center}
\begin{minted}[linenos=true]{nasm}
sub $2, $1, $3
and $12, $2, $5
or  $13, $6, $2
add $14, $2, $2
sw  $15, 100($2)
\end{minted}
\end{center}
\indent The data hazard is present in the $\$2$ register.
The $sub$ instruction will take three cycles to write to the $\$2$ register, but the $and$ and $or$ instructions are attempting to access $\$2$ before the register has a value in it.
As a result, two $nop$ instructions should be inserted after the $sub$ instruction to offset the data access.
\begin{minted}[linenos=true]{nasm}
sub $2, $1, $3
nop
nop
and $12, $2, $5
or  $13, $6, $2
add $14, $2, $2
sw  $15, 100($2)
\end{minted}
\item Identify and fix the data hazards shown below by showing the fowarding paths (hardware).
\begin{minted}[linenos=true]{nasm}
add $3, $4, $6
sub $5, $3, $2
lw  $7, 100($5)
%add $8, $7, $2
\end{minted}
There are two data hazards in this set of instructions.
In line two, register $\$3$ is being accessed by the $sub$ instruction before it is stored by the $add$ instruction.
The other hazard is with the $\$7$ register, which is used by the $add$ instruction before a value is loaded by the $lw$ instruction.\\
% \begin{tabular}{l c c c c c c c c c}
% $add$ & \sqr & & & &\\
% $sub$ & & \sqr & & &\\
% $lw$  & & & \sqr & &\\
% $nop$ & & & & \sqr &\\
% $add$ & & & &  &\sqr\\
% \end{tabular}
\item Compare the performance for the single-cycle, multi-cycle, and piplined control using the following instruction mix.
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
& Loads & Stores & Branches & Jumps & ALU\\
\hline
Instruction Mix & 25\% & 10\% & 11\% & 2\% & 52\%\\
\hline
Clock Cycles & 5 & 4 & 3 & 3 & 4\\
\hline
\end{tabular}
\end{center}
The performance for a single cycle machine is 200ps for memory access, 100ps for ALU operations, and 50ps for register read and write.
In a piplined implementation, loads, stores, and ALU operations take 1 clock cycle.
Branches take 1 clock cycle when predicted correctly, and 2 when not;
Jumps take 2 clock cycles.
\begin{enumerate}
\item What is the clock cycle time for a single-cycle data path?\\
\indent In order calculate the clock cycle time, we need to know how long each instruction takes to execute.
We can create the following table to calculate the total time:\\

@@

\item What are the average instruction times for each designs?
\indent I don't know.

@@

>>>>

\title{Computer Organization Exam Review}
\date{}

\begin{document}
\maketitle
\begin{enumerate}
\item $CPU\ Time = \frac{Clock\ Cycles}{Clock\ Rate}$
\item $CPU\ Time = Instructions \times \frac{Cycles}{Instruction} \times \frac{Seconds}{Cycle}$
\item $CPU\ Time = Clock\ Cycles \times Cycle\ Time$
\item $CPI = \frac{Clock\ Cycles}{Instructions}$
\item $CPU\ Time = Instruction\ Count \times CPI \times Cycle\ Time$
\item $CPU\ Time = Instruction\ Count \times \frac{CPI}{Clock\ Rate}$
\item $(Clock\ Cycles)_{Total} = \sum_{i = 1}^{n}\left(CPI_{i} \times Count_{i}\right)$
\item $MIPS = \frac{Instruction\ Count}{CPU\ Time \times 10^{6}}$
\item $Power = Capacitive\ Load \times (Voltage)^{2} \times Frequency$
\item $AMAT = Hit\ Time + Miss\ Rate \times Miss\ Penalty$
\end{enumerate}
\hrule
\begin{enumerate}
\item Identify the data hazards in the following code, and fix them using $nop$ (software).
\begin{center}
\begin{minted}[linenos=true]{nasm}
sub $2, $1, $3
and $12, $2, $5
or  $13, $6, $2
add $14, $2, $2
sw  $15, 100($2)
\end{minted}
\end{center}
\indent The data hazard is present in the $\$2$ register.
The $sub$ instruction will take three cycles to write to the $\$2$ register, but the $and$ and $or$ instructions are attempting to access $\$2$ before the register has a value in it.
As a result, two $nop$ instructions should be inserted after the $sub$ instruction to offset the data access.
\begin{minted}[linenos=true]{nasm}
sub $2, $1, $3
nop
nop
and $12, $2, $5
or  $13, $6, $2
add $14, $2, $2
sw  $15, 100($2)
\end{minted}
\item Identify and fix the data hazards shown below by showing the fowarding paths (hardware).
\begin{minted}[linenos=true]{nasm}
add $3, $4, $6
sub $5, $3, $2
lw  $7, 100($5)
%add $8, $7, $2
\end{minted}
There are two data hazards in this set of instructions.
In line two, register $\$3$ is being accessed by the $sub$ instruction before it is stored by the $add$ instruction.
The other hazard is with the $\$7$ register, which is used by the $add$ instruction before a value is loaded by the $lw$ instruction.\\
% \begin{tabular}{l c c c c c c c c c}
% $add$ & \sqr & & & &\\
% $sub$ & & \sqr & & &\\
% $lw$  & & & \sqr & &\\
% $nop$ & & & & \sqr &\\
% $add$ & & & &  &\sqr\\
% \end{tabular}
\item Compare the performance for the single-cycle, multi-cycle, and piplined control using the following instruction mix.
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
& Loads & Stores & Branches & Jumps & ALU\\
\hline
Instruction Mix & 25\% & 10\% & 11\% & 2\% & 52\%\\
\hline
Clock Cycles & 5 & 4 & 3 & 3 & 4\\
\hline
\end{tabular}
\end{center}
The performance for a single cycle machine is 200ps for memory access, 100ps for ALU operations, and 50ps for register read and write.
In a piplined implementation, loads, stores, and ALU operations take 1 clock cycle.
Branches take 1 clock cycle when predicted correctly, and 2 when not;
Jumps take 2 clock cycles.
\begin{enumerate}
\item What is the clock cycle time for a single-cycle data path?\\
\indent In order calculate the clock cycle time, we need to know how long each instruction takes to execute.
We can create the following table to calculate the total time:\\

@@

\item What are the average instruction times for each designs?\\
The single cycle datapath's instruction time is determined by the longest instruction time, which is the 600ps load word.
The multicycle and pipelined datapath are dependant on the longest clock cycle in the longest instruction, which is 200ps.
We then tak that and multiply it by the CPI, which yields 824ps for the multicycle datapath, and 215ps for the pipelined datapath.

@@

\item For the following diagram, explain what happens in each block for the instruction $lw\ \$t3, 100(\$t2)$.
\includegraphics[scale=0.5]{review_01}
\begin{itemize}
\item \underline{Instruction Fetch:} The machine gathers the next instruction as determined by the program counter.
The PC increments by 4 every cycle, assuming the program does not encouter any exceptions or errors.
\item \underline{Instruction Decode:} In this step, the instruction is interpreted by the machine to be the load word instruction.
\item \underline{Register Read:} The appropriate registers are parsed, and the address is decoded.
\item \underline{ALU Operation:} The ALU adds the offest to the base address, and requests the contents of that location in memory.
\item \underline {Register Write:} The machine now writes the contents of memory that have been fetched to the $\$t3$ register.
\item \underline {Instruction Fetch:} The PC is incremented by 4, and the process continues for the next instruction.
\end{itemize}
% \item A compiler designer is trying to decide between two code sequences for a particular machine.
% Based on the hardware implementation, we have the following:
% \begin{center}
% \begin{tabular}{|l|c|c|c|}
% \hline
% Class & A & B & C\\
% \hline
% CPI & 1 & 2 & 3\\
% \hline
% \end{tabular}
% \end{center}
% The first code sequence has 5 instructions:
% 2 of A, 1 of B, and 2 of C.
% The second code sequence has 6 instruction:
% 4 of A, 1 of B, and 1 of C.
% \begin{enumerate}
% \item What is the CPI for each sequence?\\
% The CPI for the first sequence is
% \[
% 2 * 1 + 1 * 2 + 2 * 3 = 10.
% \]
% The CPI for the second sequence is
% \[
% 4 * 1 + 1 * 2 + 1 * 3 = 9.
% \]
% \item Which sequence is faster, and by how much?
%
% \end{enumerate}
